\section{Introduction}
\label{sec:intro}




% \textbf{Intro}
% \begin{itemize}
%     \item Continual learning considers a general stream of experience
%     \item Question: How do we measure performance on arbitrary experience?
%     \item Background about standard regret notions. Stochastic, adversarial, ... 
%     \item These don't work in general. We need something new.
%     \item We propose ``perceived regret" as a general notion. Give definition.    
% \end{itemize}

% \textbf{Content}
% \begin{itemize}
%     \item Formal definitions and assumptions
%     \item Assumption 1,2: Upper bound on true value. Lower bound on the true value.
%     \item Assumption 3: Examiner is learning.
%     Lots of possibilities to write about. Why do we need this (vacuous upper bounds)?
%     \item Local notion of perceived regret. i.e. w.r.t. a smaller class of 
%     \item No free lunch. 
%     \item Limited capacity learner and examiner. 
%     \item Exploration and intrinsic motivation
%     \item Practical usage and experiments
%     \item 
% \end{itemize}
\wes{edit intro more}
Continual reinforcement learning imagines an agent interacting with its world and learning from its stream of experience. 
The agent chooses actions to accumulate as much reward as possible and improve its future reward. 
To this agent, the only
A questions arises: how can we measure the performance of a continual reinforcement learning agent from its stream of experience? 


\paragraph{Learning in Any World.} Our setting considers an agent interacting with a general environment. The agent takes actions $a_t$ based on its history and the environment produces rewards and observations $(r_t, o_t)$ based on the entire history.  
No assumptions are placed on the form of the environment; it is not necessarily a MDP, POMDP, etc. Thus, we allow the agent to interact with an arbitrary stream of experience $\{o_0, r_0, a_1, o_1, r_1, a_2, o_2, r_2,...\}$ with the goal of developing an evaluation procedure for intelligent agents that can 
``continue to adapt over the course of lifelong streams of experience." \citep{silver2025welcome}.


\paragraph{Motivation: Standard measures of regret are not suitable for arbitrary streams of experience.} 
Regret is traditionally defined as $V^{*} - V^{\pi_t}$ at timestep $t$ where $V^*$ represents an ``optimal" value and $V^{\pi_t}$ represents the value of the agent's current policy.
There are two classic families of regret definitions which define $V^*$ differently: 1. \textit{Expected regret}, used in stochastic bandits and MDPs. 2. \textit{Hindsight regret}, used in adversarial bandits and online learning. 

For \textit{expected regret}, $V^*$ is the value of the optimal policy, which achieves the largest expected return in the future.
This is is an unreasonable standard for the agent to achieve in a nonstationary environment since this optimal value will depend on quantities inaccessible to the agent, such as how the environment will evolve in the future. 

For \textit{hindsight regret}, $V^*$ is defined as a counterfactual: 
What if the agent had chosen past actions according to a different policy $\pi$ instead? 
Concretely, $V^*$ is the sum of rewards that \textit{would have been accumulated} by the best policy from a given set $\Pi$ if it had been run in the past. 
For example, in classic adversarial bandits, $\Pi$ is the set of policies that always choose a single action.
Then, to achieve low hindsight regret (usually termed ``external regret"), the agent has to do (almost) as well as having chosen the best fixed action in hindsight.
In nonstationary settings or MDPs, it is more difficult to define $\Pi$ without making low regret impossible to achieve \wes{add cite}.

Some works propose other notions of regret to allow for nonstationarity in the problem settings. These include adaptive regret, tracking regret, dynamic regret among others \wes{add cites}.
At a high-level, all these regret definitions limit the amount of nonstationarity to make the problem tractable, with the exact restrictions varying from one to the other. 
For this reason, these regret definitions are unsuitable for our purpose because they explicitly or implicitly place constraints on the environment. We aim to define a performance measure suitable for \textit{any} world. 

\textbf{Contributions}
In this paper, we propose the \textit{agent-examiner framework}. An \textit{examiner} $\chi$ views the same stream of experience as the agent . 

We define a novel performance measure: The \textit{perceived regret} $\rho^\chi_t$. 
This quantity is defiend

\begin{itemize}
    \item brief description
    \item We discuss assumptions on it
    \item show a no-free-lunch theorem for examiners
    \item Small experiments illustrate how perceived regret works in different, informative  problems 
    \item Discuss possible extensions
\end{itemize}



