\section{No Free Lunch for Examiners}
\label{sec:nfl}
% \begin{itemize}
%     \item What do we want in a vague sense?
%     \item Thm: Any examiner can fail. 
%     \item 
%     \item Hypothesis classes discussion? 
%     \item 
% \end{itemize}


The examiner is faced with a learning problem of its own. 
Given the stream of observations, rewards and actions, it must estimate $\Vhi$ and $\Vlow$ and produce the regret estimate $\rho^\chi_t$. 

Broadly-speaking, we desire to have an examiner $\chi$ that produces valid (in the sense of Prop. \ref{prop:valid}) estimates while keeping the bounds as tight as possible. 
Whenever the agent is learning effectively $\rho^\chi_t$ should decrease and, if the agent has an optimal policy, the examiner should output $\rho^\chi_t=0$.
Ideally, there would be a universal examiner $\chi^*$ applicable to any agent and any world while producing effective perceived regret estimates. 

Unfortunately, this is asking for too much. The situation is similar to generalization in supervised learning, where we cannot create a universal learning algorithm that generalizes well on any data distribution.
Similarly, we prove a no-free-lunch theorem showing that it is impossible to define a ``universal" examiner capable of accurately judging all agents in all worlds and discuss the implications. 

To state this theorem, we first posit some additional assumptions about the agent and environment. We assume $\Actions$ is discrete with at least two actions. We also assume the rewards are bounded in $[0, R_{max}]$ so the true values of all policies are bounded in $[0, V_{max}]$. We define some useful quantities:
Let $V_{unif}$ be the value of the uniform policy at all future steps and define $\rho_{bound} := V_{max} - V_{unif}$. This quantity is an upper bound on useful perceived regret estimates since, if the true regret is larger than $\rho_{bound}$, the agent is worse than the uniform random policy.
We say that a perceived regret $\rho$ is \textit{vacuous} if $\rho > \rho_{bound}$. 


\begin{theorem}[No Free Lunch for Examiners] 
% For any non-vacuous examiner $\chi$ and any history of experience $h \in \Hist$, there exists an environment $\mathcal{E}$ consistent with the history such that the perceived regret estimate $\rho^\chi_t$ is an invalid upper bound on the true regret.
% \wes{Could rephrase it}
The only examiner $\chi$ that produces valid perceived regret estimates (satisfying assumptions \ref{a:lb} and \ref{a:optimistic}) for any history $h \in \Hist$, any environment $\Env$ and any agent outputs a vacuous perceived regret $\rho^\chi > \rho_{bound}$  at all steps.

\end{theorem})
\begin{proof}
    We proceed by contradiction and show that, given a history $h$ and an agent, any examiner that produces a non-vacuous bound will be invalid in some environments. 
    Suppose we have an agent with history $h=\{o_0, ..., a_{t-2}, r_{t-1}, o_{t-1}\}$.
    and that the examiner $\chi$ produces a non-vacuous bound $\rho^\chi(h) < \rho_{bound}(h)$, where we make the perceived regret's dependence on the history explicit. 
    We will show that $V^{\pi} \le V_{unif}$ and $\sup_{\pi'} V^{\pi'} = V^{max}$. This will imply $V^*- V^\pi > \rho^\chi $  (perceived regret does not bound the true regret), establishing the contradiction.

  

    Given history $h$, we produce an environment $\Env_{bad}$ that generates the desired bounds. 
        
    The agent produces a distribution over its next action given its history. Let $a_{min}$ be the action with minimal probability (this exists since the action space is discrete).
    To define $\Env_{bad}$,
    we let $a_{min}$ have associated reward $\Rew(h, a_{min}) = R_{max}$ and 
    lead to a future where a unique optimal action at each step receives $R_{max}$. 
    Thus, the optimal policy has value $V_{max}$. Note that this optimal policy assigns probability 1 to $a_{min}$ for the immediate next action. 
    All other actions lead to 0 rewards forever.
    We also define $\Env_{bad}$ such that its observation function $\Tr$ leads to 
    \wes{...?}

    The uniform random policy will have associated value $V_{unif}(h)=1/|\Actions| {R_{max}} + V_{unif}(h || a_{min}, R_{max},o' )$, where $a || b$ denotes the concatenation of history $a$ followed by $b$. This expressions is due to the fact that one action has reward $R_{max}$ and other actions leads to 0 reward forever.
    Then, t

    \textit{Sketch:} The idea is that given the history $h=\{o_0, ..., a_{t-2}, r_{t-1}, o_{t-1}\}$, no matter what the agent's distribution is over the next action, there is a world where it will incur 
    . The past does not inform the future when considering an arbitrary world.
    
    agent's true value under its policy $\pi$ is $V^{\pi} \le V_{unif}$ and $\sup_{\pi'} V^{\pi'} = V^{max}$.
    This occurs if the agent's policy
        
    Thus 
    \wes{consider how to change this for stochasticity. By picking a uniform policy, you can avoid bad actions maybe? or we assume deterministic history-based policy? }
\end{proof}
\wes{high probability bound?}
\wes{add determnistic history-dependent agent}

\wes{what about randomness? Maybe forget about it for nwo}
\wes{maybe change the form of the thm}



Similar to no-free-lunch theorems in supervised learning, which tell us that over all data distributions, no model can outperform random guessing on the population loss, we find that it is impossible to produce valid non-vacuous bounds on all environments. 

where we cannot expect a learning algorithm to generalize well for all data distributions, we cannot expect the examiner to accurately assess any learner. 
The examiner must make some guess about the structure of the environment to account for possible futures and their associated rewards.  
This guess depends on the various modeling biases baked into the examiner. 
For example, the examiner may assume that the world is Markovian and produce its estimates in accordance with this assumption. This may be correct in certain environments and incorrect in others. 
The examiner's inductive bias also matters as the agent may enter previously unseen parts of the world and thus the examiner's generalization properties matter too.

Overall, we see that the examiner faces a learning problem at least as difficult as supervised learning, combining standard predictions on examples with a temporal and decision-making component. 
As such, it is expected that similar no-free-lunch theorems apply to the examiner. 
This means there are no ``universally" best examiners---we must make do with the examiners we believe are best fit for the environment at hand and cannot give general guarantees on their evaluation performance. 
Next, we discuss some choices of modeling biases we may endow the examiner with and reasonable criteria we could expect. 

% \wes{should the examiner see internals of the agent?}





\subsection{Choosing an Examiner}

% \begin{itemize}
%     \item Since universal examiners are impossible, we may have examiners that work in specific types of environments. 
%     \item Similar to supervised learning and generalization, we have to make a guess about how the world works. 
%     \item Have we pushed back the problem to the examiner? How to define learning for the examiner?
%     \item In known settings, the examiner gives meaningful bounds. 
%     \item Abstract concepts of learning: predictive information, epsilon-machine, 
%     \item Link to generalization? 
%     \item Can give some examples how to frame it...?
% \end{itemize}

Since universal examiners are impossible, we need to make tradeoffs and choose properties we may want an examiner to have. To help us decide, we may prefer examiners that are known to work in specific types of environments of interest. 
We can assess an examiner by how it behaves in certain families of environments when run on known, optimal learning agents.

For example, in stochastic bandits and finite MDPs, there are known provably-optimal learning algorithms, achieving the best (true) expected regret possible. These agents will shrink the regret per step towards zero. 
A ``good" examiner, when run on these learning agents, should also reduce its perceived regret towards zero as the agent converges towards an optimal policy. 

We may be able to create more flexible examiners, capable of dealing with larger families of environments, by considering environments where provably-optimal learning agents exist. Drawing from the vast reinforcement learning theory literature, theorists have crafted increasingly general conditions where provably low regret is achievable. For example, tabular MDPs are a special case of linear MDPs \citep{jin2020provably}, which are in turn subsumed by low Bellman rank MDPs \citep{jiang2017contextual} or low Eluder dimension \citep{russo2013eluder}. These can be  further generalized with concepts such as the Bellman-Eluder dimension \citep{jin2021bellman}, bilinear classes \citep{du2021bilinear} and many alternative concepts \citep{ayoub2020model, sun2019model}. One of the most general family of environments where low regret is possible are those with low Decision-Estimation Coefficient \citep{foster2021statistical}. 
Whenever we have a provably optimal agent for some environment family, we can use this as a benchmark to test candidate examiners. In these cases, we can require that the examiner produces a perceived regret decreasing to zero at the similar rate as the true regret. 

It is important note that, in light of the no-free-lunch theorem, it will be impossible to expand the family of environments ad infinitum. The examiner will always be unreliable in some world. 
Certain examiners may also provide invalid perceived regret, which do not bound the true regret. In a general environment, we may not have any formal guarantee, similar to generalization in supervised learning. In practice, on a unknown environment, we  forgo any formal guarantees. 
% These are all modeling decisions that a use must make. 

Without a prior knowledge of the environment, we can mitigate the possibility of poor examiner by utilizing multiple examiners in parallel on a single agent's stream of experience, resulting in several performance estimates. These can aggregated or used a distribution of assessments. Similar to ensembles, we may want to choose a diverse set of examiners, with different modeling biases and learning mechanisms. 
We can also consider pretrained models (e.g. foundation models) that have assimilated a large corpus of data. This can lead to complex inductive biases within the examiner that could potentially help it produce more useful perceived regret values. 

In the next section, we illustrate some simple experiments where the appropriate choice of examiner for an agent and environment leads to meaningful perceived regret throughout the agent's learning process.

% \textbf{Abstract notions of learning}
% An alternative approach is to posit that the examiner integrates information into its estimates. 




