We now illustrate how perceived regret behaves in practice when the examiner's inductive bias is, or is not, matched to the world. We use two multi-armed bandit testbeds chosen because the optimal learner and the true regret are both known, so $\rho^\chi_t$ can be read against ground truth. The stationary testbed has $K{=}10$ Bernoulli arms with fixed means; the switching testbed uses the same arm set but resamples the means at four equally spaced segment boundaries (dotted lines in Figure~\ref{fig:switching}). We pair each world with a \emph{good} agent (UCB1 in the stationary world, sliding-window UCB in the switching world) and a \emph{bad} agent (an undertuned $\varepsilon$-greedy, or stationarity-assuming UCB1, respectively). For each agent we run a \emph{good} examiner whose modelling assumptions match the world and a \emph{bad} examiner whose assumptions are mismatched. All curves report mean regret per step over 100 seeds with 95\% bootstrap intervals.

\textbf{Matched examiners track true regret.} In the stationary world (Figure~\ref{fig:stationary}), the good examiner's $\rho^\chi_t$ decays toward zero on the UCB1 agent, mirroring the true regret to within the shaded interval, and plateaus near the true sub-optimality gap on the $\varepsilon$-greedy agent. The bad examiner, which freezes its aspiration $\Vhi$ at a vacuous upper bound, produces a flat $\rho^\chi_t$ on the good agent and is uninformative about whether learning has occurred. In the switching world (Figure~\ref{fig:switching}), the matched SW-aware examiner produces $\rho^\chi_t$ spikes at each segment change and recovers between switches, tracking the true regret of SW-UCB; the same examiner correctly reports persistent regret for UCB1, whose stationarity assumption prevents it from re-adapting.

\textbf{Mismatched examiners fail in predictable ways.} A stationarity-assuming examiner run in the switching world (purple, Figure~\ref{fig:switching}) drifts toward zero between switches even when the underlying agent is incurring large true regret, producing an optimistic but invalid assessment. Conversely, an overly nonstationary examiner inflates $\rho^\chi_t$ on agents that have already converged, paying a constant gap to the true regret. These two regimes are the practical face of the no-free-lunch result: validity (Assumptions~\ref{a:lb}--\ref{a:optimistic}) and tightness can each be lost when the examiner's bias is incompatible with the stream.

\textbf{Procedure.} Together, these experiments suggest a simple recipe for using perceived regret on an unknown stream: (i) instantiate a small, diverse pool of examiners whose inductive biases bracket plausible world dynamics; (ii) report $\rho^\chi_t$ from each examiner alongside the agent's behaviour; and (iii) treat agreement across examiners as evidence of meaningful regret estimates, and disagreement as a diagnostic that the examiner pool, not necessarily the agent, requires revisiting.

